Address Clustering Heuristics for Account-Based Blockchain Networks: An Analysis based on a Decentraland User Set

### 2023-11-07 paper.tex
Update Section 2 and 3

Broadening the scope of entity identification, \cite{victorlüders2019} study the largest fungible token networks on Ethereum from a graph perspective. Similarly, \cite{casalebrunet2021} apply network graph analysis to various non-fungible token (NFT) ecosystems. Both find that many networks follow either a star or a hub-and-spoke pattern, similar to social network interaction graphs. Moreover, using k-means clustering, \cite{Payette2017} propose a segmentation of the Ethereum address space into four distinct behavior groups sharing similar attributes. \newline

@@

In this work, I perform entity identification on an address set containing 473,927 addresses. To obtain these addresses, a separate project collected snapshots of avatar activity in Decentraland over nine months from July 2022 to April 2023. Due to Decentraland's blockchain-based architecture, each avatar contains information about a user's Web3 address. \newline
I test the applicability of existing heuristics, and if feasible, their efficacy in detecting entities using multiple addresses is evaluated.
The main objective is to estimate the number of real-world entities these addresses represent. The clustering of addresses is limited to this set only, without considering any addresses outside of it, e.g., that were not registered in Decentraland during the given time frame. Furthermore, I introduce my heuristic for identifying address clusters and evaluate its suitability for implementation in other contexts. As of my knowledge, this work is the first to cluster addresses within a predefined set on a large scale.

The remainder of this paper is structured as follows: In Section 2, I provide a brief overview of the basic concepts necessary to understand the setup, including Ethereum accounts, tokens, Decentraland, and privacy-enhancing protocols. Section 3 describes our methodology for data collection and preparation. The fourth section provides a detailed explanation of various clustering heuristics. The fifth section applies the clustering heuristics described in section 4, after which I discuss the results of our analysis and summarize our findings in sections 6 and 7.

@@

\section{Data Collection and Preparation} 
The starting point for our work was a data set comprising 473,927 distinct Web3 addresses.
Focusing on clustering methods using on-chain data, I exclude addresses that have not been recorded on either Ethereum or Polygon. To accomplish this, information from Blockscan\footnote{\url{https://blockscan.com/}} was used. Overall, 59,651 addresses were recorded on Ethereum, 129,988 on Polygon, with 52,095 addresses appearing on both networks simultaneously. The set of ''clusterable'' addresses therefore counts 137,544 addresses. The subsets are visualized as a Venn diagram in Figure \ref{fig:Venn}.

@@

\subsection{Transactions and Token Transfer Events}
The address subsets were used to collect transaction and token transfer event data from Ethereum and Polygon. Ethereum or Polygon data can be accessed directly through a node or an application programming interface (API) provider like Infura or Etherscan. The absence of account indexing in Ethereum or Polygon poses a challenge for retrieving all past transactions and token transfers of a specific address, as it requires scanning through all blocks and token transfer events emitted by designated token contracts. Fortunately, Etherscan offers an API Endpoint for ''Accounts'' that facilitates the retrieval of transactions and token transfer events for a given address. By using our network-specific address subsets, I was able to significantly decrease the number of required API calls. 
I gathered all transactions and token transfers up until block number \texttt{17,670,000} on Ethereum (July 11, 2023) and \texttt{44,990,000} on Polygon (July 12, 2023). In total, I collected over 30 million %30,689,978
token transfer events and more than 16 million %16,092,531
 normal transactions. While the amount of transactions is similar between the two networks, Polygon has roughly three times more transfer events. \newline The output was stored in separate CSV files and imported into a MongoDB\footnote{\url{https://www.mongodb.com/}} database. The database contains distinct collections for transactions and transfers. For additional information on the data fields of each collection, see Appendix \ref{sec:database}.

\subsection{Intra-Set Asset Transfers}
\label{sec:intra-set}
I want to employ address clustering heuristics to determine the number of entities represented by the remaining 137,544 addresses. %Or, to phrase it differently, do users interact with Decentraland using multiple EOAs, and can identify addresses that belong to the same user?
\newline
To achieve this, I often only consider transactions or token transfer events where both the ''\texttt{from}'' (sender) and the ''\texttt{to}'' (recipient) address are within the address set. I call this reduced data set \textit{intra-set asset transfers}. Regarding normal transactions, this means that only native asset transfers (also called payments) are kept, as almost all addresses are EOAs. \newline
The intra-set asset transfer data set is mainly used to generate the network graphs (see Section \ref{sec:node}).
When clustering addresses using graph-based approaches, this has the advantage that I do not need to differentiate between EOA and smart contract account later on. Furthermore, it allows us to observe and visualize the asset flows between users. In total, there are 974,478 intra-set token transfers and 210,548 intra-set native asset transfers.  %101,621 Matic, 108,927 Ether

@@

Figure \ref{fig:ENS} illustrates the mechanism that allowed us to generate ground truth pairs in a simplified way: Initially, the domain name \texttt{`foo.eth`} resolves to address \texttt{`0xABC...`}. The owner has also set a reverse resolution for this address, indicated by the the edge from the address to the domain name. Next, the owner decides to change the resolved address to \texttt{`0xDEF...`}, again setting a reverse record.  As a domain can only point to one address at the time, the previous mapping from \texttt{`foo.eth`} to the old address is deleted. However, \texttt{`ABC....addr.reverse`} still points to \texttt{`foo.eth`} and querying the name corresponding to \texttt{`0xABC...`} and \texttt{`0xDEF...`}, will return \texttt{`foo.eth`}.
If this special case occurs, I assume that both addresses belong to the same entity.

@@

I applied the following methodology to find ground truth pairs: \textit{First}, I gathered all addresses that interacted with ENS to reduce the amount of name look-ups. \textit{Second}, I iterated over all addresses and checked if they reverse-resolved to an ENS domain\footnote{This was done via the ENS module within web3.py \url{https://web3py.readthedocs.io/en/stable/ens_overview.html}, which provides an interface to look up domains and addresses.}. If so, I logged the ENS domain and address in a CSV file. \textit{Third}, using our new list, I identified if two addresses referenced the same ENS domain. If matched, I recorded both addresses alongside the ENS domain. \newline
From this, I identified 11,440 reverse records and 40 pairs of addresses pointing to the same ENS name, which I utilized as our ground truth to evaluate various clustering heuristics. Although this sample is relatively small, it allows us to assess and draw conclusions about the effectiveness of the methods.

>>>>

Broadening the scope of entity identification, \cite{victorlüders2019} study the largest fungible token networks on Ethereum from a graph perspective. Similarly, \cite{casalebrunet2021} apply network graph analysis to various non-fungible token (NFT) ecosystems. Both find that many networks follow either a star or a hub-and-spoke pattern, similar to social network interaction graphs. Moreover, \cite{Payette2017} propose a partition of the Ethereum address space into four distinct behavior groups sharing similar attributes while applying k-means clustering.\newline

@@

In this work, I perform entity identification on an address set containing 473,927 addresses. To obtain these addresses, a separate project collected snapshots of avatar activity in Decentraland over nine months from July 2022 to April 2023. Due to Decentraland's blockchain-based architecture, each avatar contains information about a user's Web3 address. \newline
I test the applicability of existing heuristics, and if feasible, their efficacy in detecting entities using multiple addresses is evaluated.
The main objective is to estimate the number of real-world entities these addresses represent. The clustering of addresses is limited to this set only, without considering any addresses outside of it, e.g., that were not registered in Decentraland during the given time frame. Furthermore, I introduce a new heuristic for identifying address clusters and evaluate its suitability for implementation in other contexts. As of my knowledge, this work is the first to cluster addresses within a predefined set on a large scale.

The remainder of this paper is structured as follows: Section 2 presents a brief overview of the basic concepts necessary to understand the setup, including Ethereum accounts, tokens, Decentraland, and privacy-enhancing protocols. Section 3 describes our methodology for data collection and preparation. The fourth section provides a detailed explanation of various clustering heuristics. The fifth section applies the clustering heuristics described in section 4, after which I discuss the results of our analysis and summarize our findings in sections 6 and 7.

@@

\section{Data Collection and Preparation} 
The starting point for this work was a data set comprising 473,927 distinct Web3 addresses.
Focusing on clustering methods using on-chain data, I exclude addresses that have not been recorded on either Ethereum or Polygon. To accomplish this, I used data from Blockscan\footnote{\url{https://blockscan.com/}}. Overall, 59,651 addresses were recorded on Ethereum, 129,988 on Polygon, with 52,095 addresses appearing on both networks simultaneously. The set of ''clusterable'' addresses therefore counts 137,544 addresses. The subsets are visualized as a Venn diagram in Figure \ref{fig:Venn}.

@@

\subsection{Transactions and Token Transfer Events}
The address subsets were used to collect transaction and token transfer event data from Ethereum and Polygon. Ethereum or Polygon data can be accessed directly through a node or an application programming interface (API) provider like Infura or Etherscan. The absence of account indexing in Ethereum or Polygon poses a challenge for retrieving all past transactions and token transfers of a specific address, as it requires scanning through all blocks and token transfer events emitted by designated token contracts. Fortunately, Etherscan offers an API Endpoint for ''Accounts'' that facilitates the retrieval of transactions and token transfer events for a given address. By using the network-specific address subsets, I was able to significantly decrease the number of required API calls. 
I gathered all transactions and token transfers up until block number \texttt{17,670,000} on Ethereum (July 11, 2023) and \texttt{44,990,000} on Polygon (July 12, 2023). In total, I found over 30 million token transfer events and more than 16 million 
 normal transactions. While the amount of transactions is similar between the two networks, Polygon has roughly three times more transfer events. \newline The output was stored in separate CSV files and imported into a MongoDB\footnote{\url{https://www.mongodb.com/}} database. The database contains distinct collections for transactions and transfers. For additional information on the data fields of each collection, see Appendix \ref{sec:database}.  %30,689,978 %16,092,531

\subsection{Intra-Set Asset Transfers}
\label{sec:intra-set}
I want to employ address clustering heuristics to determine the number of entities represented by the remaining 137,544 addresses. %Or, to phrase it differently, do users interact with Decentraland using multiple EOAs, and can identify addresses that belong to the same user?
To achieve this, I often reduce the data set to only consider transactions or token transfer events where both the ''\texttt{from}'' (sender) and the ''\texttt{to}'' (recipient) address are within the address set. I call this reduced data set \textit{intra-set asset transfers}. Regarding normal transactions, this means that only native asset transfers (also referred to as payments) are kept, as almost all addresses are EOAs. \newline
The intra-set asset transfer data set is mainly used to generate the network graphs (see Section \ref{sec:node}). 
When clustering addresses through graph-based approaches, this has the advantage that contract accounts without a logical owner do not need to be excluded later on. Furthermore, it allows us to observe and visualize the asset flows between users. In total, there are 974,478 intra-set token transfers and 210,548 intra-set native asset transfers.  %101,621 Matic, 108,927 Ether

@@

Figure \ref{fig:ENS} illustrates the mechanism that allowed to generate ground truth pairs in a simplified way: Initially, the domain name \texttt{`foo.eth`} resolves to address \texttt{`0xABC...`}. The owner has also set a reverse resolution for this address, indicated by the the edge from the address to the domain name. Next, the owner decides to change the resolved address to \texttt{`0xDEF...`}, again setting a reverse record.  As a domain can only point to one address at the time, the previous mapping from \texttt{`foo.eth`} to the old address is deleted. However, \texttt{`ABC....addr.reverse`} still points to \texttt{`foo.eth`} and querying the name corresponding to \texttt{`0xABC...`} and \texttt{`0xDEF...`}, will return \texttt{`foo.eth`}.
If this special case occurs, I assume that both addresses belong to the same entity.

@@

I applied the following methodology to discover ground truth pairs: \textit{First}, I gathered all addresses that interacted with ENS to reduce the amount of name look-ups. \textit{Second}, I iterated over all addresses and checked if they reverse-resolved to an ENS domain\footnote{This was done via the ENS module within web3.py \url{https://web3py.readthedocs.io/en/stable/ens_overview.html}, which provides an interface to look up domains and addresses.}. If this was the case, the ENS domain and address were stored in a CSV file. \textit{Third}, I identified if two addresses referenced the same ENS domain. If matched, I recorded both addresses alongside the ENS domain. \newline
11,440 reverse records and 40 pairs of addresses pointing to the same name were identified and utilized as ground truth to evaluate various clustering heuristics. Although this sample is relatively small, it allows me to compare and draw conclusions about the effectiveness of the methods.
